### 2019-12-12 paper/abstract.tex
small edits

% Word limit: 250

The rotational evolution of cool dwarfs is poorly constrained after $\sim$2-3
billion years due to a lack of precise ages and rotation periods for old
main-sequence stars.
% However, recent evidence suggests that the relationship between rotation
% period and color/temperature/mass does not remain constant over large
% timescales, contradicting a long-held assumption that underlies many
% gyrochronology relations.
In this work, we use the velocity dispersions of low-mass \kepler\ dwarfs as
an age proxy, to reveal their rotational evolution and demonstrate that
kinematics could be a useful tool for calibrating gyrochronology.
% At ages less than $\sim$1 Gyr we find that a gyrochronology model, calibrated
% to the Praesepe cluster, accurately predicts the relative ages of stars.
We find that the period-\teff\ relationship of the Praesepe cluster does not
apply to stars older than around 1 Gyr.
% : a polynomial gyrochronology model, fit
% to Praesepe stars, under-predicts the ages of late K dwarfs in the field.
Although late-K dwarfs spin more slowly than early-K dwarfs when they are
young, at old ages we find that late-G and early-K dwarfs rotate at the {\it
same rate} or faster than late-K dwarfs of the same age.
% This is probably because their stronger magnetic fields lead to more efficient
% angular momentum loss and reflects the behavior of young open clusters such as
% the Pleaides, Praesepe, and the Hyades.
% However, at older ages we find that late G and early K dwarfs rotate at the
% {\it same rate} or faster than late K dwarfs of the same age.
% These oThis may be a feature of angular momentum redistribution in the stellar
% interior.
% This implies that low-mass stars lose angular momentum more slowly than
% higher-mass stars after a certain age or Rossby number.
This result agrees qualitatively with semi-empirical models that vary the rate
of surface-to-core angular momentum transport as a function of time and mass.
It also aligns with recent observations of stars in the NGC 6811 cluster,
which indicate that K dwarfs experience an epoch of stalled spin-down.
We find that the oldest \kepler\ stars with measured rotation periods are
late-K and early-M dwarfs, indicating that these stars maintain spotted
surfaces and stay magnetically active longer than more massive stars.
Finally, based on their kinematics, we find that many rapidly rotating
K-dwarfs are likely to be synchronized binaries.
% Firstly, we
% find no evidence for mass-dependent heating in a sample of K and M dwarfs in
% the \kepler\ field.

% We find that groups of stars, selected to be the same age using a
% gyrochronology relation calibrated to the Praesepe cluster, do {\it not} have
% the same velocity dispersion at all temperatures; cooler stars have greater
% velocity dispersions than hotter stars.
% We interpret this to mean that these cooler stars are older than the hotter
% ones, which implies that the period-color relation of Praesepe is not
% appropriate for old stars.
% Futhermore, our results indicate that the slope of the period-color relation
% actually {\it inverts}: at young ages early K stars rotate more rapidly than
% late K stars, but beyond $\sim$ 2 Gyr they rotate more slowly.
% These observations could be reproduced by mass-dependent dynamical heating,
% where lower-mass stars experience a greater rate of dynamical heating than
% higher-mass stars, however we find no evidence for mass-dependent heating
% within K and M dwarfs in the \kepler\ field.