### 2020-03-11 paper/conclusion.tex
incorporating Adrian's suggestions

\section{Conclusions}

In this paper, we used the \vb\ velocity dispersions of stars in the \mct\
catalog to explore the evolution of stellar rotation period as a function of
effective temperature and age.
Our conclusions are as follows:
\begin{itemize}
\item{{\bf Magnetic braking efficiency doesn't always increase with decreasing
mass for K dwarfs.}
% We found that the \citet{angus2019} relation, which is based on the
% period--color relation of the 650 Myr Praesepe cluster, does not correctly
% describe the period--age--\teff\ relation for old stars.
Although at young ages, rotation period is anti-correlated with \teff\ (as
seen in many young open clusters, include Praesepe), at intermediate ages the
relation flattens out and K dwarfs of different masses rotate at the same
rate.
At old ages, cooler K dwarfs spin more rapidly than hotter K dwarfs of the
same age.}

\item{{\bf Variable core-envelope coupling may be the cause.} We showed that
the period--\teff\ relations change shape over time in a way that
qualitatively agrees with theoretical models which include mass and
time-dependent core-envelope angular momentum transport \citep{spada2019}.}

% The period-color/\teff\ relation seems to start flattening out after $\sim$1
% Gyr (see figure \ref{fig:age_cut}), potentially around the same age, or just
% older than the period gap which is located at a gyrochronal age of around
% $\sim$1.1 Gyr (see figure \ref{fig:dispersion_period_teff}).
\item{{\bf Low-mass stars stay active longer.}
We found that the oldest stars in the \mct\ catalog are cooler than 4500
K, which suggests that lower-mass stars remain active for longer, allowing
their rotation periods to be measured at older ages.}

\item{{\bf The \kepler\ period gap may be related to core-envelope coupling.}
We speculated that the rotation period gap \citep{mcquillan2014} may separate
a young regime where stellar rotation periods decrease with increasing mass
from an old regime where periods increase with increasing mass, however more
data are needed to provide a conclusive result.
The velocity dispersions of stars increase smoothly across the rotation period
gap, indicating that the gap does not separate two distinct stellar
populations.}

\item{{\bf Rapidly rotating stars with large absolute velocities may be
synchronized binaries.}
We used kinematics to indicate that there is a population of
synchronized binaries with rotation periods less than around 10 days.}

\item{{\bf \sigmavb\ can be used as an age proxy for \kepler\ stars.} In the
    appendix of this paper we demonstrate that the dispersion of velocities in
        the direction of Galactic latitude, \vb, can be used as an age proxy,
        by showing that there is no strong evidence for mass-dependent heating
        in low-mass \kepler\ dwarfs: the velocity dispersions of K and M
        dwarfs, whose main-sequence lifetimes are longer than around 11 Gyrs,
        do not appear to increase with decreasing mass.
Although {\it vertical} velocity, \vz, is a quantity that has been
demonstrated to trace time-dependent orbital heating in the disc of the
Galaxy, most stars with measured rotation periods do not yet have radial
velocities, so we used velocity in the direction of Galactic latitude, \vb\,
as a proxy for \vz.
        Using stars in the GUMS simulation \citep{robin2012}, we showed that
        using \vb\ as a proxy for \vz\ introduces an additional
        pseudo-velocity dispersion, which increases with increasing Galactic
        latitude.
% For this reason we did not attempt to convert \vb\ dispersions into ages using
% an age-velocity dispersion relation.
However, after removing high-latitude ($b>15^\circ$) stars from the sample, we
confirmed that using \vb\ instead of \vz\ does not introduce any
mass-dependent velocity dispersion bias into the sample.
We therefore confirmed that \vb\ velocity dispersion can be used to accurately
rank stars by age.}
% , \ie\ a group of stars with a large velocity dispersion is,
% on average, older than a group of stars with a small velocity dispersion.}

\end{itemize}

% explanations for these observations, including incorrect dust corrections for
% the lowest-mass stars and an excess of companions increasing the velocity
% dispersion for these stars.

% If the period-color/\teff\ relation does invert at old ages as our results
% suggest, this would be a paradigm shift for gyrochronology.
% Stellar spin-down rate is thought to be directly tied to magnetic field
% strength, and the deeper convection zones of cooler stars generate stronger
% magnetic fields which {\it should} lead to more efficient angular momentum
% loss.
% However, the micro- and macro-physics of stellar structure and evolution and
% magnetic dynamo models are extremely complicated and a lot is still unknown
% about the magnetic behavior of stars.
% Observations like these can provide useful constraints for physical models,
% and may help to reveal new physical processes at work in stars like our own
% Sun and other planet hosts.