### 2020-03-16 paper/conclusion.tex
adding jen's small comments and a core-envelope coupling section

\section{Conclusions}

In this paper, we used the \vb\ velocity dispersions of stars in the \mct\
catalog to explore the evolution of stellar rotation period as a function of
effective temperature and age.
Our conclusions are as follows:
\begin{itemize}
\item{{\bf Spin-down rate doesn't always increase with decreasing mass for K
    dwarfs.}
Although at young ages, rotation period is anti-correlated with \teff\ (as
seen in many young open clusters, including Praesepe), at intermediate ages the
relation flattens out and K dwarfs of different masses rotate at the same
rate.
At old ages, cooler K dwarfs spin more rapidly than hotter K dwarfs of the
same age.}

\item{{\bf Variable core-envelope coupling may be the cause.} We showed that
the period--\teff\ relations change shape over time in a way that
qualitatively agrees with theoretical models which include mass and
time-dependent core-envelope angular momentum transport \citep{spada2019}.}

\item{{\bf Low-mass stars stay active longer.}
We found that the oldest stars in the \mct\ catalog are cooler than 4500
K, which agrees with previous results which show that lower-mass stars remain
        active for longer, allowing their rotation periods to be measured at
        older ages.}

\item{{\bf The \kepler\ period gap may be related to core-envelope coupling.}
We speculated that the rotation period gap \citep{mcquillan2014} may separate
a young regime where stellar rotation periods decrease with increasing mass
from an old regime where periods increase with increasing mass, however more
data are needed to provide a conclusive result.
The velocity dispersions of stars increase smoothly across the rotation period
gap, indicating that the gap does not separate two distinct stellar
populations.}

\item{{\bf Rapidly rotating stars with large absolute velocities may be
synchronized binaries.}
We used kinematics to indicate that there is a population of
synchronized binaries with rotation periods less than around 10 days.}

% \item{{\bf \sigmavb\ can be used as an age proxy for (low Galactic latitude)
%     \kepler\ stars.} In the appendix of this paper we demonstrate that the
%         dispersion of velocities in the direction of Galactic latitude, \vb,
%         can be used as an age proxy, by showing that there is no strong
%         evidence for mass-dependent heating in low-mass \kepler\ dwarfs: the
%         velocity dispersions of K and M dwarfs, whose main-sequence lifetimes
%         are longer than around 11 Gyrs, do not appear to increase with
%         decreasing mass.
% Although {\it vertical} velocity, \vz, is a quantity that has been
% demonstrated to trace time-dependent orbital heating in the disc of the
% Galaxy, most stars with measured rotation periods do not yet have radial
% velocities, so we used velocity in the direction of Galactic latitude, \vb\,
% as a proxy for \vz.
%         Using stars in the GUMS simulation \citep{robin2012}, we showed that
%         using \vb\ as a proxy for \vz\ introduces an additional
%         velocity scatter, which increases with increasing Galactic
%         latitude.
% However, after removing high-latitude ($b>15^\circ$) stars from the sample, we
% confirmed that using \vb\ instead of \vz\ does not introduce any
% mass-dependent velocity dispersion bias into the sample.
% We therefore confirmed that \vb\ velocity dispersion can be used to accurately
% rank stars by age.}

\end{itemize}